Real Time Adaptive Shape Reconstruction for Asteroid Landing

### 2018-04-03 abstract.tex
Correct error in equation
Update abstract with comments from Dr. Lee

\title{Real Time Shape Reconstruction for Near Earth Asteroid Landing}

@@

% on arrival spacecraft spend long periods mapping, and depending on mission this might be unallocable
On approach to an asteroid, spacecraft navigation and guidance is primarily based on ground measurements.
After arrival, a spacecraft will generally spend months or years in a mapping mission phase~\cite{kubota2003,cole1998}.
During this period, spacecraft sensors, such as on board optical telescopes or Light radio Detection and Ranging (LIDAR), are used to characterize the asteroid.
The resulting imagery and range data is transmitted to the ground and the resulting asteroid shape and motion is estimated. 
During this mapping phase the spacecraft must remain in a quiescent state devoted entirely to mapping the surface.
Depending on the mission type this long period of mapping is crucial to the mission, such as sample collection~\cite{gates2015}. 
However, other missions, such as asteroid mitigation, may be severely limited by the time and ground resources required to generate a surface shape.
Furthermore, the long distances involved necessitate on-board autonomy to enable to spacecraft to operate without ground communications.
The dependency on expensive ground based shape reconstruction techniques limit the ability of spacecraft to autonomously operate at asteroids.
Mitigation of this ground based surface modeling will greatly expand the range of missions possible.

@@

The radial distance of each vertex, \( v_i = \norm{\vb{v}_i}\), is assumed to be distributed according to the Gaussian distribution
\begin{align*}
    v_i \sim \mathcal{N}(r_i, w_i^2)
\end{align*}
where \( r_i \) is the initial estimate of the radial distance of vertex \( \vb{v}_i\) and \( w_i \) is the initial variance, or confidence, in the radial distance.
The radial distance of each measurement, \( p_{j,i} = \vb{p}_j\), is also assumed to be distributed according to the Gaussian distribution
\begin{align*}
    p_{j,i} \sim \mathcal{N}(r_{j,i}, w_{j,i}^2)
\end{align*}
where \( r_{j,i} = \norm{\vb{p}_{j,i}} \) defines the radial distance of the surface vector measurement and \( w_{j, i}\) defines the variance of the measurement with respect to vertex \( \vb{v}_i\).

The variance for each measurement vector is assumed to be related to the ``distance'' from the measurement to vertex \( \vb{v}_i \).
Here we use the geodesic distance to parameterize the difference, and hence  uncertainty, of associating the measurement with a given vertex.
From spherical trigonometry the central angle between measurement \( \vb{p}_i \) and vertex \( \vb{v}_i \) of the shape estimate is
\begin{align}
    \Delta \sigma_{j,i} = \arctan \parenth{\frac{\norm{\vb{p}_j \times \vb{v}_i}}{\vb{p}_j \cdot \vb{v}_i }}.
\end{align}
The variance of measurement \( \vb{p}_i \) with respect to vertex \( \vb{v}_i \) is then defined as the arc length as
\begin{align}
    w_{j, i} = \norm{\vb{p}_j} \Delta \sigma_{j,i} .
\end{align}
This approach relates the uncertainty of the measurement, \( \vb{p}_j \) with the geodesic distance to a given vertex, \( \vb{v}_i \).
As a result, measurements which are far from a vertex, where \( \Delta \sigma \) is large, will tend to have a larger variance and hence uncertainty. 
From Bayes' theorem, the posterior probability is
\begin{align}
    p(v_i | p_{j, i}) = \frac{p(p_{j, i} | v_i) p(v_i)}{p( p_{j, i})} \propto p(p_{j,i} | v_i) p(v_i).
\end{align}
From the properties of a Gaussian, the posterior probability given a measurement is also distributed according to a Gaussian distribution
\begin{align}\label{eq:posterior_probability}
    \mathcal{N} \parenth{\frac{w_{j, i}^2 r_i + w_i^2 r_{j, i}}{w_i^2 + w_{j, i}^2} , \frac{w_i^2 + w_{j, i}^2}{w_i^2  w_{j, i}^2}} .
\end{align}
From~\cref{eq:posterior_probability}, the posterior probability conditioned on the measurement is the weighted average of the prior knowledge and the measurement. 
Measurements that are far from the vertex will have a high uncertainty or variance and will have a reduced impact on the radial position of the vertex.
Additional measurements are incorporated using a weighted average of prior belief and the measurement uncertainty.

@@

\section{Preliminary Results}
\begin{figure}[h]
    \centering
    \subcaptionbox{Initial mesh estimate\label{fig:start}}{\includegraphics[height=0.2\textheight,width=0.5\textwidth,keepaspectratio]{figures/partial_initial.jpg}}~
    \subcaptionbox{First vertex added\label{fig:first}}{\includegraphics[width=0.5\textwidth]{figures/partial_0.jpg}}

    \subcaptionbox{\SI{25}{\percent} of vertices added\label{fig:quarter}}{\includegraphics[width=0.5\textwidth]{figures/partial_512.jpg}}~
    \subcaptionbox{\SI{50}{\percent} of vertices added\label{fig:half}}{\includegraphics[width=0.5\textwidth]{figures/partial_1024.jpg}}

    \subcaptionbox{\SI{75}{\percent} of vertices added\label{fig:threequarter}}{\includegraphics[width=0.5\textwidth]{figures/partial_1536.jpg}}~
    \subcaptionbox{Final reconstruction\label{fig:final}}{\includegraphics[width=0.5\textwidth]{figures/partial_2047.jpg}}
    \caption{4769 Castalia incremental reconstruction. The initial ellipsoid is incrementally modified by radially moving each vertex to best match the measurements.~\label{fig:reconstruction}}
\end{figure}
Here we present initial results of reconstructing the shape of asteroid 4769 Castalia.
Castalia is a potentially  hazardous asteroid and the first asteroid which used radar imaging to generate a 3D surface shape~\cite{hudson1994}.
In this example, we seek to incrementally build an accurate shape of Castalia from an initial triaxial ellipsoid model.
Our initial shape for Castalia is a triaxial ellipsoid with semi-axes equal to \( \SI{1.8}{\kilo\meter} \times \SI{0.8}{\kilo\meter} \times \SI{0.8}{\kilo\meter}\) and shown in~\cref{fig:start}.
We use this choice of initial estimate since the semi-major axes of an asteroid are able to be computed based on Earth-measurements~\cite{busch2011}.
Furthermore, this type of model also allows for the use of analytical gravitational models for use in preliminary analysis or in the absence of higher fidelity shape information~\cite{scheeres1994}.

>>>>

\title{Real Time Adaptive Shape Reconstruction for Asteroid Landing}

@@

% on arrival spacecraft spend long periods mapping, and depending on mission this might be unallocable
On approach to an asteroid, spacecraft navigation and guidance is primarily based on ground measurements.
After arrival, a spacecraft will generally spend months or years in a mapping mission phase~\cite{kubota2003,cole1998}.
During this period, spacecraft sensors, such as on board optical telescopes or Light radio Detection and Ranging (LIDAR), are used to characterize the asteroid.
The resulting imagery and range data is transmitted to the ground and the resulting asteroid shape and motion is estimated. 
During this mapping phase the spacecraft must remain in a quiescent state devoted entirely to mapping the surface.
Depending on the mission type this long period of mapping is crucial to the mission, such as sample collection~\cite{gates2015}. 
However, other missions, such as asteroid mitigation, may be severely limited by the time and ground resources required to generate a surface shape.
Furthermore, the long distances involved necessitate on-board autonomy to enable to spacecraft to operate without ground communications.
Similarly, during landing the spacecraft will require the ability to sense and model the surface topography in order to safely land in an unknown enviornment.
The dependency on expensive ground based shape reconstruction techniques limit the ability of spacecraft to autonomously operate at asteroids.
Mitigation of this ground based surface modeling will greatly expand the range of missions possible.

@@

The radial distance of each vertex, \( v_i = \norm{\vb{v}_i}\), is assumed to be distributed according to the Gaussian distribution
\begin{align*}
    v_i \sim \mathcal{N}(r_i, w_i^2)
\end{align*}
where \( r_i \) is the initial estimate of the radial distance of vertex \( \vb{v}_i\) and \( w_i \) is the initial variance, or confidence, in the radial distance.
The radial distance of each measurement, \( p_{j,i} = \norm{\vb{p}_j}\), is also assumed to be distributed according to the Gaussian distribution
\begin{align*}
    p_{j,i} \sim \mathcal{N}(r_{j,i}, w_{j,i}^2)
\end{align*}
where \( r_{j,i} = \norm{\vb{p}_{j,i}} \) defines the radial distance of the surface vector measurement and \( w_{j, i}\) defines the variance of the measurement with respect to vertex \( \vb{v}_i\).
Each measurement is defined by the index \( j \) while the associated vertex is defined by \( i \). 
As a result, the measurement \( p_{j, i} \) defines the distribution of measurement \( j \) with respect to vertex \( i \). 
A given measurement may be used to update one or several vertices.

The variance for each measurement vector is assumed to be related to the ``distance'' from the measurement to vertex \( \vb{v}_i \).
Here we use the geodesic distance to parameterize the difference, and hence  uncertainty, of associating the measurement with a given vertex.
From spherical trigonometry the central angle between measurement \( \vb{p}_i \) and vertex \( \vb{v}_i \) of the shape estimate is
\begin{align}
    \Delta \sigma_{j,i} = \arctan \parenth{\frac{\norm{\vb{p}_j \times \vb{v}_i}}{\vb{p}_j \cdot \vb{v}_i }}.
\end{align}
The variance of measurement \( \vb{p}_i \) with respect to vertex \( \vb{v}_i \) is then defined as the arc length as
\begin{align}
    w_{j, i} = \norm{\vb{p}_j} \Delta \sigma_{j,i} .
\end{align}
This approach relates the uncertainty of the measurement, \( \vb{p}_j \) with the geodesic distance to a given vertex, \( \vb{v}_i \).
As a result, measurements which are far from a vertex, where \( \Delta \sigma \) is large, will tend to have a larger variance and hence uncertainty. 
From Bayes' theorem, the posterior probability is
\begin{align}
    p(v_i | p_{j, i}) = \frac{p(p_{j, i} | v_i) p(v_i)}{p( p_{j, i})} \propto p(p_{j,i} | v_i) p(v_i).
\end{align}
From the properties of a Gaussian, the posterior probability given a measurement is also distributed according to a Gaussian distribution~\cite{bishop2006} and given by
\begin{align}\label{eq:posterior_probability}
    \mathcal{N} \parenth{\frac{w_{j, i}^2 r_i + w_i^2 r_{j, i}}{w_i^2 + w_{j, i}^2} , \frac{w_i^2  w_{j, i}^2}{w_i^2 +  w_{j, i}^2}} .
\end{align}
From~\cref{eq:posterior_probability}, the posterior probability conditioned on the measurement is the weighted average of the prior knowledge and the measurement. 
Measurements that are far from the vertex will have a high uncertainty or variance and will have a reduced impact on the radial position of the vertex.
Additional measurements are incorporated using a weighted average of prior belief and the measurement uncertainty.

@@

\section{Preliminary Results}
\begin{figure}[h]
    \centering
    \subcaptionbox{Initial mesh estimate\label{fig:start}}{\includegraphics[height=0.5\textheight,width=0.5\textwidth,keepaspectratio]{figures/partial_initial.jpg}}~
    \subcaptionbox{First measurement added\label{fig:first}}{\includegraphics[width=0.5\textwidth]{figures/partial_0.jpg}}

    \subcaptionbox{\SI{25}{\percent} of measurments added\label{fig:quarter}}{\includegraphics[width=0.5\textwidth]{figures/partial_512.jpg}}~
    \subcaptionbox{\SI{50}{\percent} of measurements added\label{fig:half}}{\includegraphics[width=0.5\textwidth]{figures/partial_1024.jpg}}

    \subcaptionbox{\SI{75}{\percent} of measurements added\label{fig:threequarter}}{\includegraphics[width=0.5\textwidth]{figures/partial_1536.jpg}}~
    \subcaptionbox{Final reconstruction\label{fig:final}}{\includegraphics[width=0.5\textwidth]{figures/partial_2047.jpg}}
    \caption{4769 Castalia incremental reconstruction. The initial ellipsoid is incrementally modified by radially moving each vertex to best match the measurements.~\label{fig:reconstruction}}
\end{figure}
Here we present initial results of reconstructing the shape of asteroid 4769 Castalia.
Castalia is a potentially  hazardous asteroid and the first asteroid which used radar imaging to generate a 3D surface shape~\cite{hudson1994}.
In this example, we seek to incrementally build an accurate shape of Castalia from an initial triaxial ellipsoid model.
Our initial shape for Castalia is a triaxial ellipsoid with semi-axes equal to \( \SI{1.8}{\kilo\meter} \times \SI{0.8}{\kilo\meter} \times \SI{0.8}{\kilo\meter}\) and shown in~\cref{fig:start}.
We use this choice of initial estimate since the semi-major axes of an asteroid are able to be computed based on Earth-measurements~\cite{busch2011}.
Furthermore, this type of model also allows for the use of analytical gravitational models for use in preliminary analysis or in the absence of higher fidelity shape information~\cite{scheeres1994}.
